---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  message = FALSE,
  warning = FALSE,
  dpi = 300
)
```

# ropenscipackagesreviewed

This package contains a list of R packages in the [rOpenSci registry](https://github.com/ropensci/roregistry/) and metadata on whether they have gone through the rOpenSci software review process (as of 2019-09-19).

## Installation

You can install the development version of ropenscipackagesreviewed with:

``` r
# install.packages("devtools")
devtools::install_github("sharlagelfand/ropenscipackagesreviewed)
```

## Analysis

The list of packages is available internally, in `ropenscipackagesreviewed::ropensci_packages`, or [in a CSV](https://github.com/sharlagelfand/ropenscipackagesreviewed/blob/master/data-raw/ropensci_packages.csv). The code to get the packages is [also available](https://github.com/sharlagelfand/ropenscipackagesreviewed/blob/master/data-raw/ropensci_packages.R).

If a package had a GitHub issue onboarding URL in the registry, it is considered to have gone through (or be going through) rOpenSci software review. If the issue has the ["6/approved" label](https://github.com/ropensci/software-review/issues?utf8=%E2%9C%93&q=is%3Aissue+label%3A6%2Fapproved+) on it, then I'm considering the review to be "Completed" (regardless of whether the issue is open or closed). If it doesn't, then the review is "In Progress".

```{r packages}
library(ropenscipackagesreviewed)

ropensci_packages
```

```{r n-packages}
library(dplyr)

n_ropensci_packages <- nrow(ropensci_packages)
ropensci_packages_review <- ropensci_packages %>%
  filter(software_review)
ropensci_packages_review_completed <- ropensci_packages_review %>%
  filter(software_review_status == "Completed")
ropensci_packages_review_in_progress <- ropensci_packages_review %>%
  filter(software_review_status == "In Progress")
ropensci_packages_no_review <- ropensci_packages %>%
  filter(!software_review)
```

Of the `r n_ropensci_packages` packages in the rOpenSci registry, `r scales::percent(nrow(ropensci_packages_review)/n_ropensci_packages)` (`r nrow(ropensci_packages_review)`) have completed review (`r nrow(ropensci_packages_review_completed)`) or are still in review (`r nrow(ropensci_packages_review_in_progress)`).

The following shows the number of packages that went through review each year, using the date that the "6/approved" label was added to the issue. Note that in one case (the `riem` package), [the onboarding issue](https://github.com/ropensci/software-review/issues/39) has the "6/approved" label on it, but there is no corresponding event for adding this label. In this case, the issue closed date is used.

```{r packages-reviewed-by-year}
library(ggplot2)
library(lubridate)

ropensci_packages_review_completed <- ropensci_packages_review_completed %>%
  mutate(review_year = year(review_completed))

ropensci_packages_review_completed %>%
  count(review_year) %>%
  arrange(review_year) %>%
  ggplot(aes(x = review_year, y = n)) +
  geom_col() +
  geom_text(aes(label = n, vjust = -0.5)) +
  labs(
    x = "Review Year",
    y = "Number of Packages",
    title = "rOpenSci Packages Completed Software Review, by Year"
  ) +
  theme_minimal(14)
```

We can also look at how long it takes packages to go through review. The following shows the distribution of days from when the issue was created to when the "6/approved" label was added to the issue.

```{r days-to-review}
ropensci_packages_review_completed <- ropensci_packages_review_completed %>%
  mutate(issue_created_to_review_completed = as.numeric(review_completed - issue_created) / (60 * 60 * 24))

ggplot(
  ropensci_packages_review_completed,
  aes(x = issue_created_to_review_completed)
) +
  geom_histogram(binwidth = 30) +
  scale_x_continuous(breaks = seq(0, max(ropensci_packages_review_completed[["issue_created_to_review_completed"]]), 90)) +
  labs(
    x = "Number of Days",
    y = "Number of Packages",
    title = "rOpenSci Software Review Time",
    subtitle = "Number of days from GitHub issue created to package approved."
  ) +
  theme_minimal(14)
```

```{r days-to-review-median}
issue_created_to_review_completed_summary <- summary(ropensci_packages_review_completed[["issue_created_to_review_completed"]])
```

Of all rOpenSci packages that have gone through software review, 50% complete review within `r round(issue_created_to_review_completed_summary[["Median"]], 1)` days. 

The following shows the median number of days it takes packages to go through review, by review year, along with the number of packages (shown above, but as a reminder!).

```{r days-to-review-median-by-year}
library(knitr)

ropensci_packages_review_completed %>%
  group_by(`Review Year` = review_year) %>%
  summarise(
    `Median Days from Issue Created to Review Completed` = round(median(issue_created_to_review_completed), 1),
    `Number of Packages` = n()
  ) %>%
  kable()
```
