---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%",
  message = FALSE,
  warning = FALSE,
  dpi = 300
)
```

# ropenscipackagesreviewed

This package contains a list of R packages in the [rOpenSci registry](https://github.com/ropensci/roregistry/) and metadata on whether they have gone through the rOpenSci software review process (as of 2019-09-19).

## Installation

You can install the development version of ropenscipackagesreviewed with:

``` r
# install.packages("devtools")
devtools::install_github("sharlagelfand/ropenscipackagesreviewed)
```

## Analysis

The list of packages is available internally, in `ropenscipackagesreviewed::ropensci_packages`, or [in a CSV](https://github.com/sharlagelfand/ropenscipackagesreviewed/blob/master/data-raw/ropensci_packages.csv). The code to get the packages is [also available](https://github.com/sharlagelfand/ropenscipackagesreviewed/blob/master/data-raw/ropensci_packages.R).

If a package had a GitHub issue onboarding URL in the registry, it is considered to have gone through (or be going through) rOpenSci software review. If the issue has the ["6/approved" label](https://github.com/ropensci/software-review/issues?utf8=%E2%9C%93&q=is%3Aissue+label%3A6%2Fapproved+) on it, then I'm considering the package to have been "Approved" (regardless of whether the issue is open or closed). If it doesn't, then the review is "In Progress".

```{r packages}
library(ropenscipackagesreviewed)

ropensci_packages
```

```{r n-packages}
library(dplyr)

n_ropensci_packages <- nrow(ropensci_packages)
ropensci_packages_review <- ropensci_packages %>%
  filter(software_review)
ropensci_packages_review_approved <- ropensci_packages_review %>%
  filter(review_status == "Approved")
ropensci_packages_review_in_progress <- ropensci_packages_review %>%
  filter(review_status == "In Progress")
ropensci_packages_no_review <- ropensci_packages %>%
  filter(!software_review)
```

Of the `r n_ropensci_packages` packages in the rOpenSci registry, `r scales::percent(nrow(ropensci_packages_review)/n_ropensci_packages)` (`r nrow(ropensci_packages_review)`) have gone through review (`r nrow(ropensci_packages_review_approved)`) or are still in review (`r nrow(ropensci_packages_review_in_progress)`).

The following shows the number of packages that went through review each year, using the GitHub issue *create* date for the review year. Of course, some packages may not be approved the same year the issue is opened (especially if it happens towards the end of the year), and 2019 is not over yet!

```{r packages-reviewed-by-year}
library(ggplot2)
library(lubridate)

ropensci_packages_review_approved <- ropensci_packages_review_approved %>%
  mutate(review_year = year(issue_created))

ropensci_packages_review_approved %>%
  count(review_year) %>%
  arrange(review_year) %>%
  ggplot(aes(x = review_year, y = n)) +
  geom_col() + 
  geom_text(aes(label = n, vjust = -0.5)) +
  labs(x = "Review Year",
       y = "Number of Packages",
       title = "rOpenSci Packages Completed Software Review, by Year") + 
  theme_minimal(14)
```
